---
title: "Tooth Growth Analysis"
author: "Jordan DeHerrera"
date: "November 22, 2015"
output: word_document
---

# Synopsis

The goal of the following procedures is to analyze the ToothGrowth data in R.  The analysis is organized into the following sections:
* Exploratory Data Analysis
* Confidence Intervals and Hypothesis Testing
* Assumptions

## Exploratory Data Analysis

### Necessary Libraries

```{r}
# Use ggplot2 library for exploratory graphs
library(ggplot2)
```

### Analysis

Let's look at what the data looks like with the summary function:

```{r}
# Summary of ToothGrowth data
summary(ToothGrowth)
```

Based on the summary information, we can infer the following about the data:
* There are two supplement types (OJ and VC)
* There is a dosage level with variability.  Let's find out how many unique values exist within the attribute:

```{r}
# Unique values of dose
unique(ToothGrowth$dose)
```

* Three unique dosage levels (0.5, 1.0, and 2.0)

Let's see how the population of data fits into these six different categories (3 dosages across 2 supplement types) using a boxplot graph within ggplot2:



# Confidence Intervals and Hypothesis Testing

We'll use Welch's t-test to test the null hypothesis that two distinct populations have equal means.  If the null hypothesis is rejected, then that means that 

```{r}
# Create empty variables for means
means_10 = NULL
means_1000 = NULL

# Create empty variables for distributions
expdist_10 = NULL
expdist_1000 = NULL

# Create 40 simulations with lambda of 0.2
expdist_1 = rexp(40,0.2)

# Create 10 iterations of 40 simulations
for (i in 1 : 10) expdist_10   = c(expdist_10  , rexp(40, 0.2))

# Create 1000 iterations of 40 simulations
for (i in 1 : 1000) expdist_1000 = c(expdist_1000, rexp(40, 0.2))

# Separate each of the iterations of 40 simulations
# into separate samples and then calculate the mean
# of those sample populations

# 1 iteration
means_1 <- mean(expdist_1)

# 10 iterations
means_10   <- unlist(lapply(split(expdist_10, 
              ceiling(seq_along(expdist_10)/40)),mean))

# 1000 iterations
means_1000 <- unlist(lapply(split(expdist_1000, 
              ceiling(seq_along(expdist_1000)/40)),mean))
```

# Results - Simulation Results vs Theory

## Variance

The theoretical variance of an exponential distribution is the following:

```{r}
0.2^-2
```

Here are the results of the simulations:

```{r}
var(expdist_1)
var(expdist_10)
var(expdist_1000)
```

As shown in the above, as the number of iterations increases, the variance approaches the theoretical variance.

## Distribution of the Mean

As the number of iterations increases, we can also see that the distribution begins to smooth out and appear to be closer to a normal distribution's "bell curve" type shape.  The distribution is centered around the theoretical mean which is 1.0/0.2 or 5.

```{r}
# Calculate theoretical mean
theoretical_mean <- 1.0/0.2

# Two graphs in two rows, one column
# par(mfrow = c(2,1))

# Exponential Distribution of 10 iterations
hist(means_10, main="Mean Exponential Distribution", 
     xlab="Result", 
     col="lightblue", 
     las=1, 
     breaks=10,      
     prob = TRUE)
mtext("40 simulations, 10 iterations",3)
abline(v=theoretical_mean, col="blue", lwd=3)
curve(dnorm(x, mean=mean(means_10), sd=sd(means_10)), add=TRUE, lwd=3, lty=5)

# Exponential Distribution of 1000 iterations
hist(means_1000, main="Mean Exponential Distribution", 
     xlab="Result", 
     col="lightblue", 
     las=1, 
     breaks=200, 
     prob = TRUE)
mtext("40 simulations, 1000 iterations",3)
abline(v=theoretical_mean, col="blue", lwd=3)
curve(dnorm(x, mean=mean(means_1000), sd=sd(means_1000)), add=TRUE, lwd=3, lty=5)
```